Predicting exercise activity
========================================================

The data come from http://groupware.les.inf.puc-rio.br/har.

Read training and test data from csv source files

```{r cache=TRUE}
ptrain <- read.csv("C:/Users/piisptau/Downloads/pml-training.csv", header=TRUE, stringsAsFactors=F)

ptest <- read.csv("C:/Users/piisptau/Downloads/pml-testing.csv", header=TRUE, stringsAsFactors=F)

```


Reduce the training and test sets by selecting only the following 54 columns. Columns with NA values are omitted and only numeric columns are included in the reduced set.


```{r cache=TRUE}

trainbaseset = ptrain[,c("num_window","roll_belt","pitch_belt","yaw_belt","total_accel_belt","gyros_belt_x","gyros_belt_y","gyros_belt_z","accel_belt_x","accel_belt_y","accel_belt_z","magnet_belt_x","magnet_belt_y","magnet_belt_z","roll_arm","pitch_arm","yaw_arm","total_accel_arm", "gyros_arm_x","gyros_arm_y","gyros_arm_z","accel_arm_x","accel_arm_y","accel_arm_z","magnet_arm_x","magnet_arm_y","magnet_arm_z","roll_dumbbell","pitch_dumbbell","yaw_dumbbell","total_accel_dumbbell","gyros_dumbbell_x","gyros_dumbbell_y",   "gyros_dumbbell_z", "accel_dumbbell_x","accel_dumbbell_y",  "accel_dumbbell_z",  "magnet_dumbbell_x","magnet_dumbbell_y", "magnet_dumbbell_z",  "roll_forearm","pitch_forearm","yaw_forearm",
"total_accel_forearm","gyros_forearm_x", "gyros_forearm_y","gyros_forearm_z","accel_forearm_x","accel_forearm_y",  "accel_forearm_z", "magnet_forearm_x", 
"magnet_forearm_y","magnet_forearm_z","classe")]

testing = ptest[,c("num_window","roll_belt","pitch_belt","yaw_belt","total_accel_belt","gyros_belt_x","gyros_belt_y","gyros_belt_z","accel_belt_x","accel_belt_y","accel_belt_z","magnet_belt_x","magnet_belt_y","magnet_belt_z","roll_arm","pitch_arm","yaw_arm","total_accel_arm", "gyros_arm_x","gyros_arm_y","gyros_arm_z","accel_arm_x","accel_arm_y","accel_arm_z","magnet_arm_x","magnet_arm_y","magnet_arm_z","roll_dumbbell","pitch_dumbbell","yaw_dumbbell","total_accel_dumbbell","gyros_dumbbell_x","gyros_dumbbell_y",   "gyros_dumbbell_z",  "accel_dumbbell_x","accel_dumbbell_y",  "accel_dumbbell_z",  "magnet_dumbbell_x","magnet_dumbbell_y", "magnet_dumbbell_z",  "roll_forearm","pitch_forearm","yaw_forearm",
"total_accel_forearm","gyros_forearm_x", "gyros_forearm_y", "gyros_forearm_z", "accel_forearm_x","accel_forearm_y", "accel_forearm_z", "magnet_forearm_x", 
"magnet_forearm_y","magnet_forearm_z")]


```

Divide training set into two parts, 50% of the set model training and rest for validation.


```{r cache=TRUE}

library(caret)

trainbaseset$classe = factor(trainbaseset$classe)

trainIndex = createDataPartition(y = trainbaseset$classe, p=0.5, list=FALSE) 

trainingSet = trainbaseset[trainIndex,]

validationSet = trainbaseset[-trainIndex,]

```

Add missing classe column to the test set

```{r cache=TRUE}

classe = rep("x",20)

testSet = cbind(testing, classe)

```

Make a random forest model with 4-fold cross validation

```{r cache=TRUE}

modFitRandFor <- train(trainingSet$classe ~.,data = trainingSet,method="rf", trControl=trainControl(method="cv", number=4))

modFitRandFor

```

Test the model on the validation set and get model accuracy from confusionMatrix output as 0.997. 

```{r}
library(caret)

predValValid = predict(modFitRandFor, validationSet)

confusionMatrix(validationSet$classe, predValValid)
```

Make prediction from the test set

```{r}

predictRF = predict(modFitRandFor, testSet)

predictRF

```

These test set predictions were 100% accurate.


